% !TEX root =  main.tex
%

We fix two disjoint sets $\participants = \{\ptp p, \ptp q, \ldots\}$
and $\multiroles = \{\amulti P, \amulti Q, \ldots \}$, respectively of
\emph{participants} and \emph{multiple} roles, and define the set of
\emph{roles} $\roleset = \participants \cup \multiroles$.
% \cup \multiroles^\unknownstarop \cup
% \multiroles^\unknownop$ (ranged over by $\arole, \arole_1,
% \ldots$) where $\multiroles^{\unknownstarop} = \{\unknownstar \sst
% \amulti P \in\multiroles\}$ and $\multiroles^{\unknownop} = \{
% \unknown \sst \amulti p \in \multiroles
% \}$ account for some flexibility when implementing multiroles: an
% action involving a
% $\unknownstar$ role can be optionally executed by an implementation
% of role $\amulti P$, while
% $\unknown$ establishes that exactly one implementer must execute
% that action.
%
We conventionally write multiroles with initial uppercase letter and
unique roles with initial lowercase letter.
%
%We write $\eqR$ for the least equivalence relation on $\roleset$
%satisfying
%\[\amulti P \eqR \unknownstar \hspace{2cm} \amulti P \eqR \unknown\]

Roles are to be thought of as types inhabited by instances of
processes enacting the behaviour specified in a choreography.
%
Participants are unit types while multiroles account for multiple
instances of processes all performing actions according to their role.
%
\todo{RB: forse \`e importante dire che tutti i multiruoli sono disgiunti, giusto?}
\todo{eM: direi di no; noi vorremmo che un processo possa coprire piu' ruoli}

%
Let us first define the \emph{prefixes} used in global types
with the following grammar:
% \todo{RB: ``consuming output'' mi sembra fuorviante... ma non ho proposte migliori}
%
\[\begin{array}{lcl@{\qquad\qquad}l}
  \apref & \bnfdef
  & \apref[\arole][@][@][\alocvar] & \text{(autonomous) output}
  \\ & \bnfmid
  & \apref[\arole][{}][@][\alocvar][.] & \text{(autonomous) read-only output}
  \\ & \bnfmid
  & \apref[{}][\arole][@][\alocvar] & \text{(autonomous) input}
  \\ & \bnfmid
  & \apref[{}][\arole][@][\alocvar][.] & \text{(autonomous) read}
  \\ & \bnfmid
  & \apref[\arole][\arole'][@][\alocvar] & \text{consuming interaction}
  \\ & \bnfmid
  & \apref[\arole][\arole'][@][\alocvar][.]  & \text{read-only interaction}
  \end{array}
\]
%
We syntactically distinguish two kinds of prefixes.
%
The prefixes generated by the first four productions in the grammar of
$\apref$ above are the \emph{autonomous} prefixes, that is those
prefixes that processes can execute directly on a tuple space without
coordinating with other processes; the prefixes generated by the
remaining two productions of the grammar are the \emph{interaction}
prefixes, namely those involving a role generating tuples and a role
consuming/reading them.
%
Also, tuple types are generated and accessed according to two
modalities syntactically distinguished by the round brackets around
the tuple in prefixes.
%
More precisely, when a prefix surrounds a tuple $\atuple$ with round
brackets then $\atuple$ is meant to be consumed otherwise it is meant
to be read-only.
%
We call \emph{read-only outputs} and \emph{reads}) the autonomous
read-only prefixes, outputs and inputs prefixes those prescribing the
consumption of tuples, and consuming (resp. read-only) interactions
the interations prescribing inputs (resp. reads).

Global types $\aK$ have the following syntax
\eMnote{non e' chiaro se vogliamo $\aK \parop \aK $}
\begin{eqnarray*}
  \aK & \bnfdef & \asum[@][@][][]
                  \bnfmid
                  % \aK \parop \aK \bnfmid
                  \aK \seqop \aK \bnfmid
                  X \bnfmid
                  \grec
\end{eqnarray*}
% \todo{RB: forse il simbolo $\seqop$ \`e misleading: se ho capito bene parlando con Emilio di tratta di un una specie di parallelo asimmetrico, potremmo usare $>$ o qualche simbolo simile?}
where $I$ is a finite set of indexes; we write $\nil$ for
$\asum[@][@][][]$ when $I = \emptyset$ (we omit trailing occurrences
of $\nil$) and $\apref_j.\aK_j$ instead of $\asum[@][@][][]$ when
$I = \{j\}$.

The set $\roles \apref \subseteq \roleset$ of roles in $\apref$ is
defined in the obvious way; note that $\roles \apref$ is a singleton
if, and only if, $\apref$ is an autonomous prefix.
%
Likewise, $\roles \aK$ is the set of roles in
$\aK$.

To handle recursive behaviour, the construct $\grec$ singles out a
role $\arole \in \roles \aK$ and specifies an injective function
$\aphi : \roles \aK \setminus \{\arole\} \to \locset$ such that the
locations in $\cod(\aphi)$ do not occur in $\aK$; intuitively,
$\arole$ is the participant deciding when the recursion ends and, for
all $\arole' \neq \arole$ in $\aK$, $\aphi(\arole')$ is the location
used to communicate the decision of $\arole$ to $\arole'$.
%
We omit the decorations $\arole$ and $\aphi$ when $\roles \aK = \{\arole\}$.

We extend the notions of defined and free names to global types as
follows:
\todo{RB: ma la $\nu$ pu\`o essere usata anche in una tupla di input/read (senza output)?}
\[
 \fn{\apref[\arole][@][@][\alocvar]}
 = \fn \atuple \cup \{\alocvar \mapsto \asort[loc]\} 
\qquad
 \dn{\apref[\arole][@][@][\alocvar]} 
 = \dn \atuple 
\]
omitted prefixes are defined analogously.
\[
  \begin{array}{ll}
    \begin{array}{l@{\ =\ } ll}
      \fn{\asum[@][@][][]} & \displaystyle{\bigcup_{i\in I}} \fn{\apref_i} \cup (\fn{\aK_i}\setminus\dn{\apref_i})
      \\
      \fn{\aK_1; \aK_2} 
                &
                  \fn{\aK_1}\cup\fn{\aK_2}
      \\
      \fn \alvar & (\seq \alocvar \cap \varset)
      \\
      \fn \grec & \fn {\aK} \cup \cod(\aphi)
    \end{array}
    \begin{array}{l@{\ =\ } ll}
      \dn{\asum[@][@][][]} & \displaystyle{\bigcup_{i\in I}} \dn{\apref_i} \cup \dn{\aK_i}
      \\
      \dn{\aK_1; \aK_2} 
                           &
                             \dn{\aK_1}\cup\dn{\aK_2}
      \\
      \dn \alvar & \emptyset
      \\
      \dn \grec & \dn {\aK}
    \end{array}
  \end{array}
\]
\eMnote{check fn e dn for rec}
%
We write $\names \_$ for the set of sorted names of a term, i.e.,
$\names \apref = \fn\apref \cup \dn\apref$ and similarly
$\names \aK = \fn\aK \cup \dn\aK$. A set $S$ of sorted names is
consistent, written $\consistent S$, if $x\mapsto \asort \in S$ and
$x\mapsto \asort' \in S$ implies $\asort = \asort'$.
 
The set of well-sorted terms are defined inductively as follows:

\begin{itemize}
\item $\apref$ is well-sorted if $\fn\apref\cap\dn\apref = \emptyset$ and  
$\consistent{\names\apref}$, i.e., there are no clashes/inconsistencies in the sorts of 
the names in $\atuple$ and the locality $\alocvar$
\item $\asum[@][@][][]$ is well-sorted if for all ${i\in I}$ both
  $\apref$ and ${\aK_i}$ are well-sorted and
  $\consistent{\names {\apref_i.\aK_i}}$
\item $\aK_1;\aK_2$ is well-sorted if $\aK_1$ and $\aK_2$ are
  well-sorted and $\consistent{\names{\aK_1;\aK_2}}$
\eMnote{adapt to new syntax of rec}
\item $X$ is well-sorted and $\grec$ is well-sorted if $\aK$ is
  well-sorted.
\end{itemize}


We consider terms up-to $\alpha$-renaming of defined names and
recursion variables.
%
As usual we say that a global type $\aK$ is \emph{closed} when it does
not contain free occurrences of recursion variables $X$ or free
occurrences of names, i.e., $\supp{\fn\aK} \cap \varset = \emptyset$.


\begin{example}\label{ex:cs}
  Consider the following global type that describes the interaction of
  a client $\ptp c$ with a simple service $\ptp s$ that converts
  integers into strings.
  \[
    \aK_\eqref{ex:cs} =
    \apref[\ptp c][\ptp s][{\asort[int]}][{\aloc[l]}]  \prefop
    \apref[\ptp s][\ptp c][{\asort[str]}][{\aloc[l]}]
  \]
  The client $\ptp c$ produces an integer value on the locality
  $\aloc$.
  %
  This tuple must be consumed by the server $\ptp s$, which produces
  back the converted string for the client.
\end{example}

Elaborating on the previous examples we now discuss a few features
of our setting.

\begin{example}\label{ex:CS}
  Assume that we consider client and server in \cref{ex:cs} as
  multiroles instead of single participants, and write
  \[
    \aK_\eqref{ex:CS} =
    \apref[\amulti {C}][\amulti {S}][{\asort[int]}][{\aloc[l]}] \prefop
    \apref[\amulti {S}][\amulti {C}][{\asort[str]}][{\aloc[l]}]
  \]
  In this case, $\aK_\eqref{ex:CS}$ states that each integer produced by a client
  will be consumed by a server, which will in turn produce a string
  for one of the clients.
\end{example}
The type in \cref{ex:CS} does not ensure that clients consume the
string conversion of the integer they produced, because all tuples are put at the same location $\aloc[l]$.
%
Name binders can be used to correlate tuples.
%
\begin{example}\label{ex:CSx}
  Consider
  \[
    \aK_\eqref{ex:CSx} =
    \apref[\amulti {C}][\amulti {S}][{\nu x:\asort[int]}][{\aloc[l]}] \prefop
    \apref[\amulti {S}][\amulti {C}][{x : {\asort[int]} \cdot {\asort[str]} }][{\aloc[l]}]
  \]
  $\aK_\eqref{ex:CSx}$ associates a fresh identifier $x$ to each value produced by
  a client and consumed by a server.
  %
  Despite the identifier is known only to the communicating instances,
  this does not forbid two clients to generate the same integer value.
  %
  Basically, the name $x$ allows to express constraints about the flow
  of values.
  %
  In particular, $x : \asort[int]$ in the second interaction states
  that a server must generate a tuple that contains the consumed
  integer.
  %
  Analogously, each client must consume a tuple matching the produced
  integer.
  %
  \finex
\end{example}
The choreography in \cref{ex:CSx} does not establishes a one-to-one
association between instances of $\amulti C$ and $\amulti S$.
%
In fact, an instance of $\amulti C$ not necessarily interacts with the
same instance of $\amulti S$ in the two communications when two
instances of $\amulti C$ generate the same integer in the first
interaction.
%
A one-to-one correspondence can be achieved by using fresh localities.
\begin{example}\label{ex:CSl}
  Consider
  \[
    \aK_\eqref{ex:CSl} = 
    \apref[\amulti {C}][\amulti {S}][{{\asort[int]} \cdot \nu x: {\asort[loc]}}][{\aloc[l]}]  \prefop
    \apref[\amulti {S}][\amulti {C}][{{\asort[str]}}][{x}]
  \]
  %
  A client generates a fresh locality identified by $x$, which is then
  used as the locality for the subsequent communication.
  %
  Since the name $x$ is known only to the two communicating instances,
  the second interaction can only take place between the two instances
  that know the locality $x$.
  %
  \finex
\end{example}

\todo[inline]{We think we can also deal with situations like the following, in which 
A creates a private session for B and C. 
%
\[
  \apref[\amulti {A}][\amulti {B}][{\nu y: {\asort[loc]}}][{\aloc[l]}] \prefop
\]
\[
  \apref[\amulti {A}][\amulti {C}][{y : {\asort[loc]}}][{\aloc[l]}] \prefop
\]
\[
  \apref[\amulti {B}][\amulti {C}][\asort][{y}]
\]
}

\eMnote{{definire $\grec[@][@][i]$ and fare un forward ref a ws}}
\eMnote{rimossa interazione col broker nella scelta; utile esempio per
  far vedere che non catturiamo tutti i systemi 'buoni' con wb}
\begin{example}[Auction]
\todo[inline]{ma \`e lecito usare $\nu$ in $\apref[@][\amulti {Buyer}][{{i : \asort[str]} \cdot {p : \asort[int]} \cdot {\nu l : \asort[loc]}}][{\aloc[m]}]$ o dovrebbe essere $\apref[@][\amulti {Buyer}][{{i : \asort[str]} \cdot {p : \asort[int]} \cdot {l : \asort[loc]}}][{\aloc[m]}]$?}
\[
  \begin{array}{l}
  \apref[\ptp {broker}][\amulti {Seller}][{\atuple[makeOffer]}][{\aloc[m]}][.] \prefop
  \\  	
  \grec[][][X][{\apref[\amulti {Seller}][@][{{\asort[str]} \cdot {\asort[int]} \cdot {\nu l : \asort[loc]}}][{\aloc[m]}] \prefop X}] \seqop
  \\  	
  \grec[\aphi_1][\amulti{Buyer}][Y][{\left({
      	\begin{array}{l}
          \grec[][][Z][{\apref[@][\amulti {Buyer}][{{\asort[str]} \cdot {\asort[int]} \cdot {\asort[loc]}}][{\aloc[m]}][.] \prefop Z}] \seqop
          \\
          \apref[@][\amulti {Buyer}][{{i : \asort[str]} \cdot {p : \asort[int]} \cdot {\nu l : \asort[loc]}}][{\aloc[m]}] \prefop
          \\
          \grec[\aphi_2][\amulti{Seller}][W][\left({
    	\begin{array}{l}
	\apref[\amulti {Buyer}][\amulti {Seller}][{i : \asort[{str}] \cdot {o : \asort[int]} }][{l}] \prefop
	\\
	\qquad 
	\apref[\amulti {Seller}][\amulti {Buyer}][{\asort[quit]}][{l}] \prefop
	\\
	\qquad 
	\apref[\amulti {Seller}][@][{{i : \asort[str]} \cdot {p : \asort[int]} \cdot {\nu l : \asort[loc]}}][{\aloc[m]}] \prefop
	\\
	\qquad
	Y	
	\\
	\qquad
	\chop
	\\
	\qquad 
	\apref[\amulti {Seller}][\amulti {Buyer}][{\asort[sold]}][{l}]\prefop
%	\\
%	\qquad 
%	\apref[\amulti {Seller}][\ptp {broker}][i : {\asort[str]} \cdot o : {\asort[int]} ][{\aloc[m]}]\prefop
%	\\ \qquad
	Y	
	\\
	\qquad
	\chop
	\\
	\qquad
	\apref[\amulti {Seller}][\amulti {Buyer}][{\asort[more]}][{l}] \prefop W
	\\
	\chop
	\\
        \apref[\amulti {Buyer}][\amulti {Seller}][{\asort[{noway}]}][{l}] \prefop
        \\\qquad
        \apref[\amulti {Seller}][@][{{i : \asort[str]} \cdot {p : \asort[int]} \cdot {\nu l : \asort[loc]}}][{\aloc[m]}] \prefop
          \\\qquad
          Y
	\end{array}
	}\right)]
     	\end{array}
  }\right)}]
  \end{array}
\]
where $\aphi_1 : \amulti{Seller} \mapsto \aloc_1$ and
$\aphi_2 : \amulti{Buyer} \mapsto \aloc_2$.
%
\finex
\end{example}



  
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End:
